\bf Calendar-based graphics for visualizing people's daily schedules

### 2018-10-19 main.Rmd
roughly done

The purpose of the calendar-based visualization is to provide insights into people's daily schedules relative to events such as work days, weekends, holidays, and special events. This work was originally motivated by studying foot traffic in the city of Melbourne, Australia [@ped]. There have been 43 sensors installed that count pedestrians every hour across the downtown area (Figure \ref{fig:ped-map}). The dataset can shed light on people's daily rhythms, and assist the city administration and local businesses with event planning and operational management. A routine examination of the data would involve constructing conventional time series plots to catch a glimpse of temporal patterns. The faceted plots in Figure \ref{fig:time-series-plot}, give an overall picture of the foot traffic at 3 different sensors over 2016. Further faceting by day of the week (Figure \ref{fig:facet-time}) provides a better glimpse of the daily and sub-daily pedestrian patterns.

@@

```{r ped-map, fig.width = 7, out.width = "0.7\\linewidth", fig.cap = "Map of the Melbourne city area with dots indicating sensor locations. 3 sensors have been highlighted to give a closer look at in the paper."}
```

```{r time-series-plot, fig.height = 6, fig.cap = "Time series plots showing the number of pedestrians in 2016 measured at 3 different sensors in the city of Melbourne. Colored by the sensors, small multiples of lines show that the foot traffic varies from one sensor to another in terms of both time and number. The weekly patterns look distinctive across these 3 sensors. There is an eye-catching spike which occurred at the State Library, caused by the annual White Night event on 20th of February."}
```

@@

When `polar = TRUE`, a polar transformation is carried out on the data. The computation is similar to the one described in @Wickham2012glyph. Figure \ref{fig:fs-polar} shows star plots embedded in the monthly calendar layout, which is equivalent to Figure \ref{fig:fs-2016} placed in polar coordinates.

@@

Plots can be layered. The comparison of sensors can be done by overlaying plot the values for each (Figure \ref{fig:overlay}). Differences between the pedestrian patterns at these sensors can be seen. Flagstaff Station exhibits strong commuters patterns, with fewer pedestrian counts during the weekends and public holidays. This suggests that Flagstaff Station has limited functionality on non-work days. From Figure \ref{fig:overlay} it can be seen that the State Library has a similar temporal trend to Flagstaff Station on non-work days. The nighttime events, such as White Night and New Year's Eve, have barely affected the operation of Flagstaff Station but heavily affected the incoming and outgoing traffic to Flagstaff Station and the State Library.

```{r overlay, fig.height = 7, fig.cap = "Overlaying line graphs of the 3 sensors in the monthly calendar. Flagstaff station is not as busy as the other two on non-work days."}
```

@@

```{r boxplot, fig.height = 6.5, fig.cap = "Side-by-side boxplots of hourly counts for all the 43 sensors in December 2016, with the loess smooth line superimposed on each day. It shows the hourly distribution in the city as a whole. There is one sensor attracting a larger number of people on New Year's Eve than the rest."}
```

<!-- ```{r chn, dev.args = list(family = "wqy-microhei"), fig.height = 6.5, fig.cap = "(ref:chn-cap)"} -->
<!-- ``` -->

>>>>

The purpose of the calendar-based visualization is to provide insights into people's daily schedules relative to events such as work days, weekends, holidays, and special events. This work was originally motivated by studying foot traffic in the city of Melbourne, Australia [@ped]. There have been 43 sensors installed that count pedestrians every hour across the downtown area (Figure \ref{fig:ped-map}). The dataset can shed light on people's daily rhythms, and assist the city administration and local businesses with event planning and operational management. A routine examination of the data would involve constructing conventional time series plots to catch a glimpse of temporal patterns. The faceted plots in Figure \ref{fig:time-series-plot}, give an overall picture of the foot traffic at three different sensors over 2016. Further faceting by day of the week (Figure \ref{fig:facet-time}) provides a better glimpse of the daily and sub-daily pedestrian patterns.

@@

```{r ped-map, fig.width = 7, out.width = "0.7\\linewidth", fig.cap = "Map of the Melbourne city area with dots indicating sensor locations. These three highlighted sensors will be inspected in the paper: (1) the State Library--a public library, (2) Flagstaff Station--a train station, closed on non-work days, (3) Birrarung Marr: an outdoor park hosting many cultural and sports events."}
```

```{r time-series-plot, fig.height = 6, fig.cap = "Time series plots showing the number of pedestrians in 2016 measured at three different sensors in the city of Melbourne. Colored by the sensors, small multiples of lines show that the foot traffic varies from one sensor to another in terms of both time and number. The weekly patterns look distinctive across these three sensors. There are eye-catching spikes which occurred at the State Library and Birrarung Marr, caused by the annual White Night event on 20th of February."}
```

@@

When `polar = TRUE`, a polar transformation is carried out on the data. The computation is similar to the one described in @Wickham2012glyph. Figure \ref{fig:fs-polar} shows star plots embedded in the monthly calendar layout, which is equivalent to Figure \ref{fig:fs-2016} placed in polar coordinates. The bimodal workday shape is also visible as boomerangs.

@@

Plots can be layered. The comparison of sensors can be done by overlaying plot the values for each (Figure \ref{fig:overlay}). Differences between the pedestrian patterns at these sensors can be seen. Flagstaff Station exhibits strong commuters patterns, with fewer pedestrian counts during the weekends and public holidays. This suggests that Flagstaff Station has limited functionality on non-work days. From Figure \ref{fig:overlay} it can be seen that Birrarung Marr has a distinct temporal trend from the other two all year around. The nighttime events, such as White Night and New Year's Eve, have barely affected the operation of Flagstaff Station but heavily affected the incoming and outgoing traffic to the State Library and Birrarung Marr.

```{r overlay, fig.height = 7, fig.cap = "Overlaying line graphs of the three sensors in the monthly calendar. Three sensors demonstrate very different traffic patterns. Birrarung Marr tends to attract many pedestrians for special events held on weekends, by contrasting to the bimodal commuting traffic at Flagstaff Station."}
```

@@

```{r boxplot, fig.height = 8, fig.cap = "Side-by-side boxplots of hourly counts for all the 43 sensors in December 2016, with the loess smooth line superimposed on each day. It shows the hourly distribution in the city as a whole. There is one sensor attracting a larger number of people on New Year's Eve than the rest."}
```

### 2018-10-19 main.tex
roughly done

The purpose of the calendar-based visualization is to provide insights
into people's daily schedules relative to events such as work days,
weekends, holidays, and special events. This work was originally
motivated by studying foot traffic in the city of Melbourne, Australia
\citep{ped}. There have been 43 sensors installed that count pedestrians
every hour across the downtown area (Figure \ref{fig:ped-map}). The
dataset can shed light on people's daily rhythms, and assist the city
administration and local businesses with event planning and operational
management. A routine examination of the data would involve constructing
conventional time series plots to catch a glimpse of temporal patterns.
The faceted plots in Figure \ref{fig:time-series-plot}, give an overall
picture of the foot traffic at 3 different sensors over 2016. Further
faceting by day of the week (Figure \ref{fig:facet-time}) provides a
better glimpse of the daily and sub-daily pedestrian patterns.

@@

\caption{Map of the Melbourne city area with dots indicating sensor locations. 3 sensors have been highlighted to give a closer look at in the paper.}\label{fig:ped-map}
\end{figure}

@@

\caption{Time series plots showing the number of pedestrians in 2016 measured at 3 different sensors in the city of Melbourne. Colored by the sensors, small multiples of lines show that the foot traffic varies from one sensor to another in terms of both time and number. The weekly patterns look distinctive across these 3 sensors. There is an eye-catching spike which occurred at the State Library, caused by the annual White Night event on 20th of February.}\label{fig:time-series-plot}
\end{figure}

@@

When \texttt{polar\ =\ TRUE}, a polar transformation is carried out on
the data. The computation is similar to the one described in
\citet{Wickham2012glyph}. Figure \ref{fig:fs-polar} shows star plots
embedded in the monthly calendar layout, which is equivalent to Figure
\ref{fig:fs-2016} placed in polar coordinates.

@@

Plots can be layered. The comparison of sensors can be done by
overlaying plot the values for each (Figure \ref{fig:overlay}).
Differences between the pedestrian patterns at these sensors can be
seen. Flagstaff Station exhibits strong commuters patterns, with fewer
pedestrian counts during the weekends and public holidays. This suggests
that Flagstaff Station has limited functionality on non-work days. From
Figure \ref{fig:overlay} it can be seen that the State Library has a
similar temporal trend to Flagstaff Station on non-work days. The
nighttime events, such as White Night and New Year's Eve, have barely
affected the operation of Flagstaff Station but heavily affected the
incoming and outgoing traffic to Flagstaff Station and the State
Library.

@@

\caption{Overlaying line graphs of the 3 sensors in the monthly calendar. Flagstaff station is not as busy as the other two on non-work days.}\label{fig:overlay}
\end{figure}

>>>>

The purpose of the calendar-based visualization is to provide insights
into people's daily schedules relative to events such as work days,
weekends, holidays, and special events. This work was originally
motivated by studying foot traffic in the city of Melbourne, Australia
\citep{ped}. There have been 43 sensors installed that count pedestrians
every hour across the downtown area (Figure \ref{fig:ped-map}). The
dataset can shed light on people's daily rhythms, and assist the city
administration and local businesses with event planning and operational
management. A routine examination of the data would involve constructing
conventional time series plots to catch a glimpse of temporal patterns.
The faceted plots in Figure \ref{fig:time-series-plot}, give an overall
picture of the foot traffic at three different sensors over 2016.
Further faceting by day of the week (Figure \ref{fig:facet-time})
provides a better glimpse of the daily and sub-daily pedestrian
patterns.

@@

\caption{Map of the Melbourne city area with dots indicating sensor locations. These three highlighted sensors will be inspected in the paper: (1) the State Library--a public library, (2) Flagstaff Station--a train station, closed on non-work days, (3) Birrarung Marr: an outdoor park hosting many cultural and sports events.}\label{fig:ped-map}
\end{figure}

@@

\caption{Time series plots showing the number of pedestrians in 2016 measured at three different sensors in the city of Melbourne. Colored by the sensors, small multiples of lines show that the foot traffic varies from one sensor to another in terms of both time and number. The weekly patterns look distinctive across these three sensors. There are eye-catching spikes which occurred at the State Library and Birrarung Marr, caused by the annual White Night event on 20th of February.}\label{fig:time-series-plot}
\end{figure}

@@

When \texttt{polar\ =\ TRUE}, a polar transformation is carried out on
the data. The computation is similar to the one described in
\citet{Wickham2012glyph}. Figure \ref{fig:fs-polar} shows star plots
embedded in the monthly calendar layout, which is equivalent to Figure
\ref{fig:fs-2016} placed in polar coordinates. The bimodal workday shape
is also visible as boomerangs.

@@

Plots can be layered. The comparison of sensors can be done by
overlaying plot the values for each (Figure \ref{fig:overlay}).
Differences between the pedestrian patterns at these sensors can be
seen. Flagstaff Station exhibits strong commuters patterns, with fewer
pedestrian counts during the weekends and public holidays. This suggests
that Flagstaff Station has limited functionality on non-work days. From
Figure \ref{fig:overlay} it can be seen that Birrarung Marr has a
distinct temporal trend from the other two all year around. The
nighttime events, such as White Night and New Year's Eve, have barely
affected the operation of Flagstaff Station but heavily affected the
incoming and outgoing traffic to the State Library and Birrarung Marr.

@@

\caption{Overlaying line graphs of the three sensors in the monthly calendar. Three sensors demonstrate very different traffic patterns. Birrarung Marr tends to attract many pedestrians for special events held on weekends, by contrasting to the bimodal commuting traffic at Flagstaff Station.}\label{fig:overlay}
\end{figure}
